Shallow EDSLs and Object-Oriented Programming

### 2017-04-18 ICFP2017/response.md

We thank all the reviewers for their constructive comments.

Review A
===

- if you support multiple interpretations, then how do I write DSL terms, without committing to a particular interpretation of the DSL?

There are different options to build terms:

1) An easy way that offers some weak form of modularity but not independent from multiple interpretations is to provide smart constructors of same signatures in different versions of DSLs. For example,

```
object Circuit1 {
  def fan(x: Int) = new Fan1 {val n=x}
  def beside(x: Circuit1, y: Circuit1) = new Beside1 {val c1=x; val c2=y}
  ...
}

object Circuit2 {
  def fan(x: Int) = new Fan2 {val n=x}
  def beside(x: Circuit2, y: Circuit2) = new Beside2 {val c1=x; val c2=y}
  ...
}
```
By switching import statements from `Circuit1` to `Circuit2`, we can have a circuit that supports different interpretations without touching the construction code

2) Object Algebras are complementary to our own techniques.
By building on the strength of Object Algebras, we can have terms that are independent from interpretations.
Here is the Object Algebra interface for the circuit DSL:

```
trait CircuitFactory[C] {
  def fan(x: Int): C
  def beside(x: C, y: C): C
  ...
}
```

This interface can be used to construct interpretation-independent terms:

```
def mkCircuit[C](f: CircuitFactory[C]): C = f.beside(f.fan(2),f.fan(2))
```

Now, smart constructor definitions are moved to a concrete factory:

```
object Circuit1Factory extends CircuitFactory[Circuit1] {
  def fan(x: Int) = new Fan1 {val n=x}
  def beside(x: Circuit1, y: Circuit1) = new Beside1 {val c1=x; val c2=y}
  ...
}
object Circuit2Factory extends CircuitFactory[Circuit2] {
  def fan(x: Int) = new Fan2 {val n=x}
  def beside(x: Circuit1, y: Circuit1) = new Beside2 {val c1=x; val c2=y}
  ...
}
```

Similar concrete factories can be defined for `Circuit2`.
By supplying different concrete factories to `mkCircuit`, we obtain circuits that support different  interpretations.

Review B
===

Review C
===

- "Scala inheritance is used to save a tiny bit of redefinition"
inheritance allows reuse of existing interpretations
subtyping allows new interpretations

- This paper strangely (to me) spends 2 sentences acknowledging that the approach is a bad fit for DSLs that are used with program transformation -- as if that were not a crucial feature for many DSL domains, including the most realistic one treated in this paper!

- The OOP style demonstrated here forces a linearization of the method signatures -- each new widget inherits from a particular old one, even if the two widgets in question define two methods that aren't mutually recursive.

No. For example, `Circuit2` shown on page 6 can be defined separately:

```
trait Circuit2 {
  def depth: Int
}
trait Identity2 extends Circuit2 {
  def depth = 0
}
...
```

We can merge `Circuit1` and `Circuit2` via multiple trait inheritance:

```
trait Circuit extends Circuit1 with Circuit2
trait Identity extends Circuit12 with Idenity1 with Identity2
...
```

### 2017-05-02 JFP/Main.tex
JFP format

\documentclass[acmlarge, anonymous, review]{acmart}\settopmatter{printfolios=true}

\usepackage{xspace}
\usepackage{url}
\usepackage{syntax}
\newcommand{\dsl}{\textsc{Scans}\xspace}
\newcommand{\interp}{\textsc{Interpreter}\xspace}
%\newcommand{\an}[3]{{\color{#2} {\sc #1}:#3}}
\newcommand{\an}[3]{}
\newcommand{\weixin}[1]{\an{weixin}{cyan}{#1}}
\newcommand{\bruno}[1]{\an{bruno}{blue}{#1}}

@@

\makeatletter\if@ACM@journal\makeatother
%% Journal information (used by PACMPL format)
%% Supplied to authors by publisher for camera-ready submission
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmYear{2017}
\acmMonth{1}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\else\makeatother
%% Conference information (used by SIGPLAN proceedings format)
%% Supplied to authors by publisher for camera-ready submission
\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
\acmYear{2017}
\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\fi

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission
\setcopyright{none}             %% For review submission
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

@@

% Title portion
\title{Shallow EDSLs and Object-Oriented Programming}
\subtitle{Functional Pearl}
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}
}
\email{first1.last1@inst1.edu}          %% \email is recommended

@@

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}

\keywords{keyword1, keyword2, keyword3}

\maketitle

>>>>

\documentclass{jfp1}

\usepackage{xspace}
\usepackage{url}
\usepackage{syntax}
\usepackage{textcomp}
\newcommand{\dsl}{\textsc{Scans}\xspace}
\newcommand{\interp}{\textsc{Interpreter}\xspace}
%\newcommand{\an}[3]{{\color{#2} {\sc #1}:#3}}
\newcommand{\an}[3]{}
\newcommand{\weixin}[1]{\an{weixin}{cyan}{#1}}
\newcommand{\bruno}[1]{\an{bruno}{blue}{#1}}

@@

%% Bibliography style
\bibliographystyle{jfp}

@@

% Title portion
\title{Shallow EDSLs and Object-Oriented Programming}

\author[W. Zhang and B. Oliveira]
        {WEIXIN ZHANG and BRUNO C. D. S. OLIVEIRA\\
         The University of Hong Kong, Hong Kong}

\maketitle[f]

@@
